Points Above Replacement: A New NBA Metric to Evaluate Player Performance

### 2022-04-05 Manuscript/Manuscript.tex
somm comments

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}

@@

\usepackage[labelfont={footnotesize,bf} , textfont=footnotesize]{caption}
\captionsetup{labelformat=simple, labelsep=period}
\newcommand\num{\addtocounter{equation}{1}\tag{\theequation}}
\renewcommand{\theequation}{\arabic{equation}}
\makeatletter
\renewcommand\tagform@[1]{\maketag@@@ {\ignorespaces {\footnotesize{\textbf{Equation}}} #1.\unskip \@@italiccorr }}
\makeatother
\setlength{\intextsep}{10pt}
\setlength{\abovecaptionskip}{2pt}
\setlength{\belowcaptionskip}{-10pt}
\renewcommand{\textfraction}{0.10}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\floatpagefraction}{0.90}

@@

\newenvironment{comment}%
{\begin{quoting}\noindent\small\it\ignorespaces%
  }{\end{quoting}}
 \title{Points Above Replacement: A New NBA Metric to Evaluate Player Performance}
\author{
Brian Krikorian*, Jun Yan \\ \medskip 
Department of Statistics, University of Connecticut, Storrs, CT  \\  \medskip 
Students: brian.krikorian@uconn.edu* \\
Mentor: jun.yan@uconn.edu 
}

@@

   \end{center}
\end{titlepage}
% Abstracts are required.
\section*{Abstract}
In attempting to predict the score of an NBA game, players who are injured, suspended, or for any other 
reason not playing have a significant impact on said prediction. Baseball uses a statistic called Wins Above 
Replacement, or WAR, to tell how many wins a team would gain or lose if they replaced a certain player 
with one who is league average. The goal of this paper is to create a similar statistic, called Points Above 
Replacement, or PAR, that uses statistics from throughout the season to predict how many points a team 
gains or loses by missing a certain player, and thus having to replace him with one who is "league 
average". The statistic is standardized and adjusted by both position and value to a player's given team. 
The model roughly follows the calculation of WAR in baseball, assigning values to points created on offense 
and defense, as well as a positional and team adjustment. A comparison with the created metric along with 
other well-known advanced player metrics is done via a multiple regression.

@@

\vfill
\begin{table}[H]
  \caption{Coefficients for Each Statistic in Offensive and Defensive Points}
  \label{tab:Coefficientstable}
\centering
\begin{tabular}[t]{lccllll}
  \toprule
  Statistic & Point Guard & Shooting Guard & Small Forward & Power Forward & Center\\
  \midrule
 Effective FG Percentage & 200 & 200 & 200 & 200 & 200\\
 \midrule
 3 Point Percentage & 300 & 275 & 250 & 225 & 200\\
 \midrule
 Free Throw Percentage & 100 & 90 & 80 & 70 & 60\\
 \midrule
 Offensive Rebounds & 1 & 1.5 & 2 & 2.5 & 3\\
 \midrule
 Assists & 4 & 3.5 & 3 & 2.5 & 2\\
 \midrule
 Points & 2 & 2 & 2 & 2 & 2\\
 \midrule
 Turnovers & -5 & -4.5 & -4 & -3.5 & -3\\
 \midrule
 Defensive Rebounds & 2 & 2.5 & 3 & 3.5 & 4\\
 \midrule
 Total Rebounds & 3 & 3.25 & 3.5 & 3.75 & 4\\
 \midrule
 Steals & 5 & 4.5 & 4 & 3.5 & 3\\
 \midrule
 Blocks & 3 & 3.5 & 4 & 4.5 & 5\\
 \midrule
 Personal Fouls & -3 & -3 & -3 & -3 & -3\\
  \bottomrule
\end{tabular}
\end{table}

@@

Once all supporting statistics of PAR have been created, all that is left to do is combine them to achieve our PAR values for each player in the NBA. The formula for PAR is as follows:

\begin{equation}
((Offensive Points + DefensivePoints) * (Team Adjustment))/(Points Per Win)
\end{equation}

This gives a good idea of how many points a player accounts for during a game, how important that is for 
his given team, and how important it is compared to all other players. The last step is to standardize this 
PAR statistic, so that the mean is 0, and the standard deviation is 1. Logically, this helps the idea of the 
metric in the sense that a perfectly league average player should have a PAR of 0, and his presence or 
absence from his team neither adds nor subtracts value. Thus, we have created a new NBA player metric, 
Points Above Replacement, PAR.

>>>>

\documentclass[12pt, titlepage]{article}
\usepackage[margin=1in]{geometry}

@@

\usepackage[labelfont={footnotesize,bf} , textfont=footnotesize]{caption}
\captionsetup{labelformat=simple, labelsep=period}
% \newcommand\num{\addtocounter{equation}{1}\tag{\theequation}}
% \renewcommand{\theequation}{\arabic{equation}}
\makeatletter
\renewcommand\tagform@[1]{\maketag@@@ {\ignorespaces {\footnotesize{\textbf{Equation}}} #1.\unskip \@@italiccorr }}
\makeatother
\setlength{\intextsep}{10pt}
\setlength{\abovecaptionskip}{2pt}
\setlength{\belowcaptionskip}{-10pt}
\renewcommand{\textfraction}{0.10}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\floatpagefraction}{0.90}

@@

\newenvironment{comment}%
{\begin{quoting}\noindent\small\it\ignorespaces%
  }{\end{quoting}}

\title{Points Above Replacement: A New NBA Metric to Evaluate Player Performance}
\author{
Brian Krikorian*, Jun Yan \\ \medskip 
Department of Statistics, University of Connecticut, Storrs, CT  \\  \medskip 
Students: brian.krikorian@uconn.edu* \\
Mentor: jun.yan@uconn.edu 
}

@@

   \end{center}
 \end{titlepage}

% Abstracts are required.

\jy{watch linewidth in source; keep it under 80}

\section*{Abstract}
In attempting to predict the score of an NBA game, players who are injured, suspended, or for any other 
reason not playing have a significant impact on said prediction. Baseball uses a statistic called Wins Above 
Replacement, or WAR, to tell how many wins a team would gain or lose if they replaced a certain player 
with one who is league average. The goal of this paper is to create a similar statistic, called Points Above 
Replacement, or PAR, that uses statistics from throughout the season to predict how many points a team 
gains or loses by missing a certain player, and thus having to replace him with one who is "league 
average". The statistic is standardized and adjusted by both position and value to a player's given team. 
The model roughly follows the calculation of WAR in baseball, assigning values to points created on offense 
and defense, as well as a positional and team adjustment. A comparison with the created metric along with 
other well-known advanced player metrics is done via a multiple regression.

@@

\jy{no hard coded spacing like vfill}

\jy{no horizontal lines in tables; float with tbp}

\begin{table}[tbp]
  \caption{Coefficients for Each Statistic in Offensive and Defensive Points}
  \label{tab:Coefficientstable}
\centering
\begin{tabular}[t]{lccllll}
  \toprule
  Statistic & Point Guard & Shooting Guard & Small Forward & Power Forward & Center\\
  \midrule
 Effective FG Percentage & 200 & 200 & 200 & 200 & 200\\
 \midrule
 3 Point Percentage & 300 & 275 & 250 & 225 & 200\\
 \midrule
 Free Throw Percentage & 100 & 90 & 80 & 70 & 60\\
 \midrule
 Offensive Rebounds & 1 & 1.5 & 2 & 2.5 & 3\\
 \midrule
 Assists & 4 & 3.5 & 3 & 2.5 & 2\\
 \midrule
 Points & 2 & 2 & 2 & 2 & 2\\
 \midrule
 Turnovers & -5 & -4.5 & -4 & -3.5 & -3\\
 \midrule
 Defensive Rebounds & 2 & 2.5 & 3 & 3.5 & 4\\
 \midrule
 Total Rebounds & 3 & 3.25 & 3.5 & 3.75 & 4\\
 \midrule
 Steals & 5 & 4.5 & 4 & 3.5 & 3\\
 \midrule
 Blocks & 3 & 3.5 & 4 & 4.5 & 5\\
 \midrule
 Personal Fouls & -3 & -3 & -3 & -3 & -3\\
  \bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:Coefficientstable} summarizes ...

@@

Once all supporting statistics of PAR have been created, all that is left to do is combine them to achieve our PAR values for each player in the NBA. The formula for PAR is
\begin{equation}
  \label{eq:par}
  \frac{(\text{Offensive Points} + \text{Defensive Points}) \times \text{Team Adjustment}}
  {\text{Points Per Win}}.
\end{equation}
This gives a good idea of how many points a player accounts for during a game, how important that is for 
his given team, and how important it is compared to all other players. The last step is to standardize this 
PAR statistic, so that the mean is 0, and the standard deviation is 1. Logically, this helps the idea of the 
metric in the sense that a perfectly league average player should have a PAR of 0, and his presence or 
absence from his team neither adds nor subtracts value. Thus, we have created a new NBA player metric, 
Points Above Replacement, PAR.
